Timeliness of security updates for Android

### 2014-08-06 securityupdates.tex
Add api_gpcomp_rdiff plot which clearly shows how we have early adopters and so are conservative
Remove user agent data section as we don't need it
Prod abstract and intro

\begin{abstract}
Android is the most popular smartphone platform and seeks to provide better security than popular desktop operating systems through increased compartmentalisation and a permission system for apps.
This security relies on the absence of root privilege exploits which allow apps to bypass protections.
Many such exploits have been found and fixed.
Our hypothesis is that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
We used Device Analyzer data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average over \daMeanInsecurityPerc\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPerc\ of devices run the most recent version of Android.
There was also a period of several months when no devices ran secure versions of Android.
We compare different manufacturers and operators to find which are best at promptly distributing security updates and examine how to improve the situation. %TODO operators analysis
Many phones are sold on long (18 month) contracts with monthly payments yet stop receiving updates before the end of the contract, while Windows XP could be purchased for a one off payment in October 2001 and received security updates until April 2014.
\end{abstract}

\section{Introduction}
Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are at least \daNumManufacturers\ manufacturers which make devices which run Android.
Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least \daNumNetworkOperators) make further modifications.
Hence when Google produces an update to Android, the update may have to pass through the manufacturer and operator before reaching the user.
The manufacturer and operator have little financial incentive to perform this work as they have already sold the device.
Operators are particularly cautious because if they ship a broken update then they will get many support requests which is expensive.
However there is ongoing legal action to force operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.
CESG, which advises the UK government on how to secure its computer systems, recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.
We provide this information in~\S\ref{sec:security_scoring}.

@@

The data from Device Analyzer (DA) we want to use to investigate the proportion of devices exposed to different vulnerabilities is the OS version.
Unfortunately there is no authoritative source of OS version information and so we cannot check directly whether our data is representative.
However Google has published API version information every month since December 2009 and we have collated this information.\footnote{\url{https://www.cl.cam.ac.uk/~drt24/android/versiondistribution/}}
While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the DA data on API versions is similar to the Google Data on API versions then the DA data on OS versions should be representative.
Figure \ref{fig:play_api} shows the data from Google and Figure \ref{fig:da_api} shows the data from Device Analyzer and they appear similar.
Figure \ref{fig:da_gp_comp_diff} shows the difference between Figures \ref{fig:play_api} and \ref{fig:da_api}.
It shows that the Device Analyzer data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions.
This means that the OS version information from Device Analyzer is likely to be overestimating the prevalence of new OS versions and hence our results are a conservative estimate of the security of Android.
%TODO we want a statistical metric to claim this strongly with.
\begin{figure}
 \centering
 \includegraphics[width=\columnwidth]{figures/googleplayapi}
 \caption{Google Play data on proportion of devices running different Android API versions}
 \label{fig:play_api}
\end{figure}
\begin{figure}
 \centering
 \includegraphics[width=\columnwidth]{figures/norm_api_gpcomp}
 \caption{Device Analyzer data on proportion of devices running different Android API versions}
 \label{fig:da_api}
\end{figure}
\begin{figure}
 \centering
 \includegraphics[width=\columnwidth]{figures/api_gpcomp_diff}
 \caption{Difference between Device Analyzer and Google Play data on the proportion of devices running different Android API versions}
 \label{fig:da_gp_comp_diff}
\end{figure}

@@

\subsection{User-Agent data}
TODO

>>>>

\begin{abstract}
Android is the most popular smartphone platform and seeks to provide better security than popular desktop operating systems through increased compartmentalisation and a permission system for apps.
This security relies on the absence of root privilege exploits which allow apps to bypass protections.
Many such exploits have been found and fixed.
Our hypothesis is that these fixes do not promptly reach the users devices and so much of the time users are running versions of Android known to be vulnerable.
We used Device Analyzer data from over \daOSYearsOfData\ years and \daNumOSDevices\ devices and found that on average over \daMeanInsecurityPerc\ of devices were exposed to known root privilege vulnerabilities and only \daUpdatednessPerc\ of devices run the most recent version of Android.
There was also a period of several months when no devices ran secure versions of Android.
We compare different manufacturers and operators to find which are best at promptly distributing security updates and examine how to improve the situation. %TODO operators analysis
\end{abstract}

\section{Introduction}
Android has \percMarketShare\ of the smartphone market and while the core development is controlled by Google there are at least \daNumManufacturers\footnote{From the Device Analyzer data} manufacturers which make devices which run Android.
Android receives regular updates  which add new features and fix bugs, including security vulnerabilities.
Many of these manufacturers customise the version of Android they ship and sometimes network operators (of which there are at least \daNumNetworkOperators) make further modifications.
Hence when Google produces an update to Android, the update may have to pass through the manufacturer and operator before reaching the user.
The manufacturer and operator have little financial incentive to perform this work as they have already sold the device.
Operators are cautious because if they ship a broken update then they will get many support requests which is expensive.
However there is ongoing legal action to force operators to ship updates for security vulnerabilities~\cite{Soghoian2013}.
Many phones are sold on long (18 month) contracts with monthly payments yet stop receiving updates before the end of the contract, while Windows XP could be purchased for a one off payment in October 2001 and received security updates until April 2014.
CESG, which advises the UK government on how to secure its computer systems, recommends picking Android devices from manufactures which are good at shipping security updates promptly~\cite{CESG2013} but it does not state which manufacturers are better.
We provide this information in~\S\ref{sec:security_scoring}.

@@

The data from Device Analyzer (DA) we want to use to investigate the proportion of devices exposed to different vulnerabilities is the OS version.
Unfortunately there is no authoritative source of OS version information and so we cannot check directly whether our data is representative.
However Google has published API version information every month since December 2009 and we have collated this information.\footnote{\url{https://www.cl.cam.ac.uk/~drt24/android/versiondistribution/}}
While API versions are too coarse grained to use for security update detection they are closely related to OS versions and so if the DA data on API versions is similar to the Google Data on API versions then the DA data on OS versions should be representative.
Figure \ref{fig:play_api} shows the data from Google and Figure \ref{fig:da_api} shows the data from Device Analyzer and they appear similar.
Figure \ref{fig:da_gp_comp_diff} shows the difference between Figures \ref{fig:play_api} and \ref{fig:da_api}, normalising for days since the API version was released.
It shows that the Device Analyzer data systematically overestimates the prevalence of new API versions and underestimates the prevalence of old API versions.
This means that the OS version information from Device Analyzer is likely to be overestimating the prevalence of new OS versions and hence our results are a conservative estimate of the security of Android.
%TODO we want a statistical metric to claim this strongly with.
\begin{figure}
 \centering
 \includegraphics[width=\columnwidth]{figures/googleplayapi}
 \caption{Google Play data on proportion of devices running different Android API versions}
 \label{fig:play_api}
\end{figure}
\begin{figure}
 \centering
 \includegraphics[width=\columnwidth]{figures/norm_api_gpcomp}
 \caption{Device Analyzer data on proportion of devices running different Android API versions}
 \label{fig:da_api}
\end{figure}
\begin{figure}
 \centering
 \includegraphics[width=\columnwidth]{figures/api_gpcomp_rdiff}
 \caption{Difference between Device Analyzer and Google Play data on the proportion of devices running different Android API versions}
 \label{fig:da_gp_comp_diff}
\end{figure}

@@
